---
title: "Data Manipulation For Model"
author: "Molly Sauter"
date: "2022-11-17"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
library(tidyverse)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

## Clean Raw Data

Read in the raw data and create clean variables for the flu type.
```{r types}
flu_dat <- read.csv("PHIRST Nasopharyngeal specimens 2016-2018 Flu 2022-05-03 (1).csv")
flu_dat$TYPE <- NA
flu_dat$CT <- NA
flu_dat <- flu_dat %>%
  select(indid, funum, fuid, npsdatecol, TYPE, CT, everything()) 

# Types
flu_dat$TYPE <- ifelse(flu_dat$npsinfa == "1" & flu_dat$npsh3 == "1" 
                      & flu_dat$npsinfb == "0" & flu_dat$npsh1 == "0", "AH3", flu_dat$TYPE)
flu_dat$TYPE <- ifelse(flu_dat$npsinfa == "1" & flu_dat$npsh1 == "1" 
                      & flu_dat$npsinfb == "0" & flu_dat$npsh3 == "0", "AH1", flu_dat$TYPE)
flu_dat$TYPE <- ifelse(flu_dat$npsinfb == "1" & flu_dat$npsyam == "1" 
                      & flu_dat$npsinfa == "0" & flu_dat$npsvic == "0", "BYam", flu_dat$TYPE)
flu_dat$TYPE <- ifelse(flu_dat$npsinfb == "1" & flu_dat$npsvic == "1" 
                      & flu_dat$npsinfa == "0" & flu_dat$npsyam == "0", "BVic", flu_dat$TYPE)
flu_dat$TYPE <- ifelse((flu_dat$npsinfa == "1" & flu_dat$npsh1 == "0" &  flu_dat$npsh3 == "0") &
                        flu_dat$npsinfb == "0" 
                      | (flu_dat$npsinfa == "1" & is.na(flu_dat$npsh1) & is.na(flu_dat$npsh3)), 
                      "ANoSub", flu_dat$TYPE)
flu_dat$TYPE <- ifelse((flu_dat$npsinfb == "1" & flu_dat$npsinfa == "0" & flu_dat$npsyam == "0" &  flu_dat$npsvic == "0") 
                      | (flu_dat$npsinfb == "1" & flu_dat$npsinfa == "0" & is.na(flu_dat$npsyam) & is.na(flu_dat$npsvic)), 
                      "BNoSub", flu_dat$TYPE)

# Variable to denote unsubtyped sample
flu_dat[ ,'unsub_sam'] <- NA 
flu_dat$unsub_sam <- ifelse(flu_dat$TYPE == "ANoSub", 1, flu_dat$unsub_sam)
flu_dat$unsub_sam <- ifelse(flu_dat$TYPE == "BNoSub", 1, flu_dat$unsub_sam)

# Check ANoSub
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "ANoSub" & lag(flu_dat$TYPE, n=1) != "ANoSub" 
                      & (lag(flu_dat$npsinfb, n=1) != 1) & !is.na(lag(flu_dat$TYPE, n=1)), 
                      lag(flu_dat$TYPE, n=1), flu_dat$TYPE) 
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "ANoSub" & lag(flu_dat$TYPE, n=2) != "ANoSub" 
                      & (lag(flu_dat$npsinfb, n=2) != 1) & !is.na(lag(flu_dat$TYPE, n=2)), 
                      lag(flu_dat$TYPE, n=2), flu_dat$TYPE) 
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "ANoSub" & lead(flu_dat$TYPE, n=1) != "ANoSub" 
                      & (lead(flu_dat$npsinfb, n=1) != 1) & !is.na(lead(flu_dat$TYPE, n=1)), 
                      lead(flu_dat$TYPE, n=1), flu_dat$TYPE) 
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "ANoSub" & lead(flu_dat$TYPE, n=2) != "ANoSub" 
                      & (lead(flu_dat$npsinfb, n=2) != 1) & !is.na(lead(flu_dat$TYPE, n=2)), 
                      lead(flu_dat$TYPE, n=2), flu_dat$TYPE) 

# Check BNoSub
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "BNoSub" & lag(flu_dat$TYPE, n=1) != "BNoSub" 
                      & (lag(flu_dat$npsinfa, n=1) != 1) & !is.na(lag(flu_dat$TYPE, n=1)), 
                      lag(flu_dat$TYPE, n=1), flu_dat$TYPE) 
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "BNoSub" & lag(flu_dat$TYPE, n=2) != "BNoSub" 
                      & (lag(flu_dat$npsinfa, n=2) != 1) & !is.na(lag(flu_dat$TYPE, n=2)), 
                      lag(flu_dat$TYPE, n=2), flu_dat$TYPE) 
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "BNoSub" & lead(flu_dat$TYPE, n=1) != "BNoSub" 
                      & (lead(flu_dat$npsinfa, n=1) != 1) & !is.na(lead(flu_dat$TYPE, n=1)), 
                      lead(flu_dat$TYPE, n=1), flu_dat$TYPE) 
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "BNoSub" & lead(flu_dat$TYPE, n=2) != "BNoSub" 
                      & (lead(flu_dat$npsinfa, n=2) != 1) & !is.na(lead(flu_dat$TYPE, n=2)), 
                      lead(flu_dat$TYPE, n=2), flu_dat$TYPE) 

# Coinfection with AH3BVic -- only 1, it occurs in Agincourt in 2016
flu_dat$TYPE <- ifelse(flu_dat$npsinfa == "1" & flu_dat$npsh3 == "1" & flu_dat$npsinfb == "1" 
                      & (flu_dat$npsyam == "0" | is.na(flu_dat$npsyam)), "AH3BVic", flu_dat$TYPE)
# Coinfection with AH3BYam - 2, occur in Klerksdorp, in 2017
flu_dat$TYPE <- ifelse(flu_dat$npsinfa == "1" & flu_dat$npsh3 == "1" & flu_dat$npsinfb == "1" 
                      & flu_dat$npsyam == "1", "AH3BYam", flu_dat$TYPE)
# Coinfection with AH3H1 - 22, occurred in both 2016 and 2017, both sites
flu_dat$TYPE <- ifelse(flu_dat$npsinfa == "1" & flu_dat$npsh3 == "1" & flu_dat$npsh1 == "1" 
                      & !is.na(flu_dat$npsh3) & !is.na(flu_dat$npsh1), 
                      "AH3H1", flu_dat$TYPE)
# Coinfection with BYamVic - 2, occurred in 2016, in both sites
flu_dat$TYPE <- ifelse(flu_dat$npsinfb == "1" & flu_dat$npsyam == "1" & flu_dat$npsvic == "1"
                      & !is.na(flu_dat$npsyam) & !is.na(flu_dat$npsvic), 
                      "BYamVic", flu_dat$TYPE)

flu_dat$TYPE <- ifelse(flu_dat$flu == "Neg", "Negative", flu_dat$TYPE)
flu_dat$TYPE <- ifelse(flu_dat$flu == "", "Missing", flu_dat$TYPE)

```

Add in individual profile information and clean variables where necessary.
```{r profile}
primary_l <- c(unique(flu_dat$indid))
profile <- read.csv('PHIRST Individual 2016-2018 2022-05-03.csv') %>%
  rename(indid = ind_id)%>% 
  rename(age = age_cat_at_consent) %>%
  filter(indid %in% primary_l)
flu_dat <- flu_dat %>%
  left_join(profile, by = "indid") 

# Clean date
flu_dat$date <- NA  
flu_dat$npsdatecol <- as.Date(flu_dat$npsdatecol, format = "%d/%m/%Y")
flu_dat$date <- ifelse(flu_dat$year == 2016, as.integer(flu_dat$npsdatecol - as.Date("2016-05-02")), flu_dat$date)
flu_dat$date <- ifelse(flu_dat$year == 2017, as.integer(flu_dat$npsdatecol - as.Date("2017-01-16")), flu_dat$date)
flu_dat$date <- ifelse(flu_dat$year == 2018, as.integer(flu_dat$npsdatecol - as.Date("2018-01-15")), flu_dat$date)

# Clean age categories
flu_dat$age_cat <- NA 
flu_dat$age_cat <- ifelse(flu_dat$age <= 4, "<5", flu_dat$age_cat)
flu_dat$age_cat <- ifelse(flu_dat$age >= 5 & flu_dat$age <= 11, "5-11", flu_dat$age_cat)
flu_dat$age_cat <- ifelse(flu_dat$age >= 12 & flu_dat$age <= 18, "12-18", flu_dat$age_cat)
flu_dat$age_cat <- ifelse(flu_dat$age >= 19 & flu_dat$age <= 40, "19-40", flu_dat$age_cat)
flu_dat$age_cat <- ifelse(flu_dat$age >= 41, ">40", flu_dat$age_cat)

# Clean HIV variable 1) for positive/negative 2) for immune suppression as measured by 
flu_dat$HIV <- NA
flu_dat$HIV <- ifelse(flu_dat$arv_current_vl == "Positive" | flu_dat$arv_current_vl == "Yes", "HIV Positive", flu_dat$HIV)
flu_dat$HIV <- ifelse(flu_dat$arv_current_vl == "HIV negative" | flu_dat$arv_current_vl == "Negative" | flu_dat$arv_current_vl == "No", "HIV Negative", flu_dat$HIV)
flu_dat$HIV <- ifelse(flu_dat$arv_current_vl == "HIV status not known" | flu_dat$arv_current_vl == "Unknown", "HIV Unknown", flu_dat$HIV)

flu_dat$suppression_clean <- ifelse(flu_dat$suppression_clean == "Unknown", "No viral load results available", flu_dat$suppression_clean)
flu_dat$suppression_clean <- ifelse(flu_dat$suppression_clean == "Yes", "Suppressed throughout", flu_dat$suppression_clean)
flu_dat$suppression_clean <- ifelse(flu_dat$suppression_clean == "No", "Never suppressed", flu_dat$suppression_clean)
flu_dat$suppression_clean <- ifelse(flu_dat$suppression_clean == "Unknown", "No viral load results available", flu_dat$suppression_clean)

# Clean BMI --- need to clean up data first

flu_dat <- flu_dat %>%
  select(indid, fuid, npsdatecol, date, year, sex, TYPE, CT, site, age, age_cat, HIV, suppression_clean, everything())

rm(profile, primary_l)
```

## Dealing with un-subtyped samples. 
First assign type based on whether or not there is a positively subtyped sample within two samples (done in first section), then assign the rest based on the dominate strain circulating during that year. (Un-subtyped samples of type A in 2016 are excluded since their subtype could not be reliably assumed.) Once clustered, this effectively means that unsubtyped samples would be subtyped if there is a positive subtype sample within 14 days. 
```{r unsub}
# Here assign further based on dominate circulating strain
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "ANoSub" & flu_dat$year == 2017, "AH3", flu_dat$TYPE)
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "ANoSub" & flu_dat$year == 2018, "AH1", flu_dat$TYPE)
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "BNoSub" & flu_dat$year == 2016, "BVic", flu_dat$TYPE)
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "BNoSub" & flu_dat$year == 2017, "BYam", flu_dat$TYPE)
flu_dat$TYPE <- ifelse(flu_dat$TYPE == "BNoSub" & flu_dat$year == 2018, "BVic", flu_dat$TYPE)

# Remove TYPE A un-subtyped from 2016 
flu_dat <- flu_dat %>%
  filter(TYPE != "ANoSub")
```

## Co-Infections
Create a label for co-infections and then create new observations for them. 
```{r coinfections}

co_dat <- flu_dat %>% 
  filter(TYPE == "AH3H1" | TYPE == "AH3BVic" | TYPE == "BYamVic" | TYPE == "AH3BYam")
for (i in 1:nrow(co_dat)) {
  co_dat <- co_dat %>%
    rbind(co_dat[i,])
}
co_dat <- co_dat %>%
  arrange(fuid) %>%
  # also going to add identifier to the new fuid
  mutate(fuid_co = fuid)

# go by each type of coinfection
h3h1 <- co_dat %>%
  filter(TYPE == "AH3H1")
h3h1$TYPE <- ifelse(h3h1$fuid_co == lead(h3h1$fuid_co), "AH3", h3h1$TYPE) 
h3h1$TYPE <- ifelse(h3h1$fuid_co == lag(h3h1$fuid_co) & !is.na(lag(h3h1$fuid_co)), "AH1", h3h1$TYPE)
h3h1$fuid_co <- ifelse(h3h1$fuid_co == lag(h3h1$fuid_co) & !is.na(lag(h3h1$fuid_co)), paste(h3h1$fuid_co, "-1"), h3h1$fuid_co)

byamvic <- co_dat %>%
  filter(TYPE == "BYamVic")
byamvic$TYPE <- ifelse(byamvic$fuid_co == lead(byamvic$fuid_co), "BYam", byamvic$TYPE)
byamvic$TYPE <- ifelse(byamvic$fuid_co == lag(byamvic$fuid_co) & !is.na(lag(byamvic$fuid_co)), "BVic", byamvic$TYPE)
byamvic$fuid_co <- ifelse(byamvic$fuid_co == lag(byamvic$fuid_co) & !is.na(lag(byamvic$fuid_co)), paste(byamvic$fuid_co, "-1"), byamvic$fuid_co)

h3bns <- co_dat %>%
  filter(TYPE == "AH3BVic")
h3bns$TYPE <- ifelse(h3bns$fuid_co == lead(h3bns$fuid_co), "AH3", h3bns$TYPE)
h3bns$TYPE <- ifelse(h3bns$fuid_co == lag(h3bns$fuid_co) & !is.na(lag(h3bns$fuid_co)), "BVic", h3bns$TYPE)
h3bns$fuid_co <- ifelse(h3bns$fuid_co == lag(h3bns$fuid_co) & !is.na(lag(h3bns$fuid_co)), paste(h3bns$fuid_co, "-1"), h3bns$fuid_co)

h3byam <- co_dat %>%
  filter(TYPE == "AH3BYam")
h3byam$TYPE <- ifelse(h3byam$fuid_co == lead(h3byam$fuid_co), "AH3", h3byam$TYPE)
h3byam$TYPE <- ifelse(h3byam$fuid_co == lag(h3byam$fuid_co) & !is.na(lag(h3byam$fuid_co)), "BYam", h3byam$TYPE)
h3byam$fuid_co <- ifelse(h3byam$fuid_co == lag(h3byam$fuid_co) & !is.na(lag(h3byam$fuid_co)), paste(h3byam$fuid_co, "-1"), h3byam$fuid_co)

coinf_refine <- h3h1 %>%
  rbind(byamvic, h3bns, h3byam) %>%
  # create variable to note the co-infection is occurring
  mutate(coinfect = 1)

# combine with other data to make refined data 
non_coinf <- flu_dat %>%
  filter(TYPE == "Missing" | TYPE == "Negative" | TYPE == "AH1" | TYPE == "AH3" |TYPE == "BYam" | TYPE == "BVic" | TYPE == "ANoSub" | TYPE == "BNoSub") %>%
  mutate(coinfect = 0,fuid_co = fuid) 

flu_dat <- non_coinf %>%
  rbind(coinf_refine) %>%
  arrange(fuid) %>%
  select(1:2, fuid_co, 3:9, coinfect, everything())

rm(byamvic, co_dat, coinf_refine, h3bns, h3byam, h3h1, non_coinf, i)
```

## Cluster samples for infection
Clustering algorithm to pair together samples in an infection, with new infections defined by the presence of a new strain or two weeks of negative samples since the previous positive samples. 
```{r cluster}
clus <- flu_dat %>%
  filter(flu == "Pos")

clus$type_num <- NA
clus$type_num <- ifelse(clus$TYPE == "AH3", 0, clus$type_num)
clus$type_num <- ifelse(clus$TYPE == "AH1", 15, clus$type_num)
clus$type_num <- ifelse(clus$TYPE == "BYam", 30, clus$type_num)
clus$type_num <- ifelse(clus$TYPE == "BVic", 45, clus$type_num)

indid_l <- unique(c(clus$indid))
clusters <- c()

for (n in indid_l) {
  #subset data for that individual
  sub <- clus %>%
    filter(indid == n) 
  if (nrow(sub) == 1) {
    clusters <- c(clusters, 1)
  }
  if (nrow(sub) > 1) {
    sub <- sub %>%
      arrange(fuid) %>%
      select(date, type_num)
    sub_dist <- dist(sub, method = "euclidean")
    hc_sub <- hclust(sub_dist, method = "single")
    
    clusters <- c(clusters, cutree(hc_sub, h = 14))
  }
}

clus$infcluster <- clusters
clus <- clus %>% 
  # making new variable to combine indid and the cluster/infection number (easier to work with later)
  mutate(indid_inf = paste(indid, "_I", infcluster, sep = "")) %>%
  select(- type_num)
neg <- flu_dat %>%
  filter(flu != "Pos") %>%
  mutate(infcluster = NA) %>% 
  mutate(indid_inf = NA)
flu_dat <- clus %>%
  rbind(neg) %>%
  arrange(fuid) %>%
  select(1:7, infcluster, indid_inf, everything())

rm(clus, hc_sub, neg, sub, clusters, indid_l, n, sub_dist)
```

## Reinfections
Want to create one variable to denote if reinfection is of the same type and another to denote 
```{r reinfections}

# make a new data frame that will only contain the first event of each of the infections
one_dat <- data.frame()
indid_inf <- unique(flu_dat$indid_inf[which(!is.na(flu_dat$indid_inf))])
for (n in indid_inf) {
  sub <- flu_dat %>%
    filter(indid_inf == n)
  one_dat <- rbind(one_dat, sub[1, ])
}

# make variable for matching type and matching subtype
# can manually count up occurrences due to coinfections and subtract those out
one_dat$sametype <- NA

for(i in 1:nrow(one_dat)) {
  if(one_dat$infcluster[i] == 1) {
    next()
  }
  # if second instance then compare to just one prior
  if(one_dat$infcluster[i] == 2 & one_dat$flutp[i] == one_dat$flutp[i - 1]) {
    one_dat$sametype[i] <- 1 
  }
  if(one_dat$infcluster[i] == 2 & is.na(one_dat$sametype[i])) {
    one_dat$sametype[i] <- 0 
  }

  # if third instance then compare to both prior
  if(one_dat$infcluster[i] == 3 & one_dat$flutp[i] == one_dat$flutp[i - 1] &
    one_dat$flutp[i] == one_dat$flutp[i-2]) {
    one_dat$sametype[i] <- 1 
  }
  if(one_dat$infcluster[i] == 3 & is.na(one_dat$sametype[i])) {
    one_dat$sametype[i] <- 0 
  }

  # if fourth instance then compare to three prior
  if(one_dat$infcluster[i] == 4 & one_dat$flutp[i] == one_dat$flutp[i - 1] & 
     one_dat$flutp[i] == one_dat$flutp[i-2] & one_dat$flutp[i] == one_dat$flutp[i-3]) {
    one_dat$sametype[i] <- 1 
  }
  if(one_dat$infcluster[i] == 4 & is.na(one_dat$sametype[i])) {
    one_dat$sametype[i] <- 0 
  }

  # if fifth instance then compare to four prior
  if(one_dat$infcluster[i] == 5 & one_dat$flutp[i] == one_dat$flutp[i - 1] &
     one_dat$flutp[i] == one_dat$flutp[i-2] & one_dat$flutp[i] == one_dat$flutp[i-3] &
     one_dat$flutp[i] == one_dat$flutp[i-4]) {
    one_dat$sametype[i] <- 1 
  }
  if(one_dat$infcluster[i] == 5 & is.na(one_dat$sametype[i])) {
    one_dat$sametype[i] <- 0 
  }
}

# recombine with the rest of the data
ind <- c(unique(one_dat$fuid_co))
all_dat <- flu_dat %>%
  filter(!(fuid_co %in% ind)) %>%
  mutate(sametype = NA)

flu_dat <- all_dat %>%
  rbind(one_dat) %>%
  arrange(fuid) %>%
  select(1:9, sametype, everything())

rm(all_dat, i, ind)
```

## CT Value
Using the Ct value from the detection assay so that all come from shared probe with a shared limit of detection.
```{r}
# CT
flu_dat$CT <- ifelse(flu_dat$TYPE == "AH3" | flu_dat$TYPE == "AH1", flu_dat$npsinfact, flu_dat$CT)
flu_dat$CT <- ifelse(flu_dat$TYPE == "BYam" | flu_dat$TYPE == "BVic", flu_dat$npsinfbct, flu_dat$CT)
flu_dat$CT <- ifelse(flu_dat$TYPE == "Negative", 37, flu_dat$CT)
```

## Reducing data frame for model

Removing individuals who do not have any infection episodes. Excluding infection episodes that have two missing episodes before or after the start or end of the episode, and episodes that are cutoff by the start or the end of a collection period. Remove samples that are missing. 
```{r exclusions}

# Going to make second data frame for model with exclusions
model_dat <- flu_dat

# Incomplete episodes
model_dat$start <- NA 
model_dat$end <- NA 
indid_inf <- unique(model_dat$indid_inf[which(!is.na(model_dat$indid_inf))])
new_dat <- data.frame()
for(n in indid_inf){
  sub <- model_dat %>%
    filter(indid_inf == n)
  sub$start[1] <- 1
  sub$end[nrow(sub)] <- 1
  new_dat <- rbind(new_dat, sub)
} 
neg <- model_dat %>%
  filter(TYPE == "Missing" | TYPE == "Negative")
new_dat <- rbind(new_dat, neg) %>%
  arrange(fuid)
start_miss <- c(unique(new_dat$indid_inf[which(new_dat$start == 1 & lag(new_dat$TYPE) == "Missing" & lag(new_dat$TYPE, n = 2) == "Missing")])) # 19 of these episode
end_miss <- c(unique(new_dat$indid_inf[which(new_dat$end == 1 & lead(new_dat$TYPE) == "Missing" & lead(new_dat$TYPE, n = 2) == "Missing")])) # 25 of these episodes 
end_indid <- c(unique(new_dat$indid_inf[which(new_dat$end == 1 & new_dat$indid != lead(new_dat$indid))])) # 20 of these episodes
exclude <- unique(c(start_miss, end_miss, end_indid)) # 62 total episodes to be excluded 
model_dat <- model_dat %>%
  filter(!(indid_inf %in% exclude))

# Remove individuals who do not have any infection episode or only had an episode previously excluded
infect <- c(unique(model_dat$indid[which(model_dat$infcluster >=1)]))
model_dat <- model_dat %>%
  filter(indid %in% infect)

# remove 
rm(neg, new_dat, sub, end_indid, end_miss, exclude, indid_inf, start_miss, n, infect)

# Missing samples 
model_dat <- model_dat %>%
  filter(TYPE != "Missing")

```

## Clean up infection episode data for model
Need to create a two week bubble around the positive samples of an infection for the model. For co-infections this requires doubling data points. 
```{r}
# want to first double for those with 2 (either coinfection or reinfection)
two <- c(unique(model_dat$indid[which(model_dat$infcluster == 2)]))
for (i in two) {
  sub <- model_dat %>%
    filter(indid == i)
  neg <- sub %>%
    filter(is.na(indid_inf)) %>%
    mutate(indid_inf = paste(i, "_I", 2, sep = ""))
  sub$indid_inf <- ifelse(is.na(sub$indid_inf), paste(i, "_I", 1, sep = ""), sub$indid_inf)
  model_dat <- model_dat %>%
    filter(indid != i) %>%
    rbind(sub, neg) %>%
    arrange(fuid)
  rm(sub, neg)
}
rm(i, two)

# Make another if there is a third infection
three <- c(unique(model_dat$indid[which(model_dat$infcluster == 3)]))
for (i in three) {
  add <- model_dat %>%
    filter(indid == i & TYPE == "Negative" & indid_inf == paste(i, "_I", 1, sep = "")) %>%
    mutate(indid_inf = paste(i, "_I", 3, sep = ""))
  model_dat <- model_dat %>%
    rbind(add) %>%
    arrange(fuid)
  rm(add)
}
rm(i, three)

# for four
four <- c(unique(model_dat$indid[which(model_dat$infcluster == 4)]))
for (i in four) {
  add <- model_dat %>%
    filter(indid == i & TYPE == "Negative" & indid_inf == paste(i, "_I", 1, sep = "")) %>%
    mutate(indid_inf = paste(i, "_I", 4, sep = ""))
  model_dat <- model_dat %>%
    rbind(add) %>%
    arrange(fuid)
  rm(add)
}
rm(i, four)

# for five
five <- c(unique(model_dat$indid[which(model_dat$infcluster == 5)]))
for (i in five) {
  add <- model_dat %>%
    filter(indid == i & TYPE == "Negative" & indid_inf == paste(i, "_I", 1, sep = "")) %>%
    mutate(indid_inf = paste(i, "_I", 5, sep = ""))
  model_dat <- model_dat %>%
    rbind(add) %>%
    arrange(fuid)
  rm(add)
}
rm(i, five)

# And now give infection labels for remaining 
one <- c(unique(model_dat$indid[which(is.na(model_dat$indid_inf))]))
for (i in one) {
  sub <- model_dat %>%
    filter(indid == i) %>% 
    mutate(indid_inf = paste(i, "_I", 1, sep = ""))
  model_dat <- model_dat %>%
    filter(indid != i  ) %>%
    rbind(sub) %>% 
    arrange(fuid)
  rm(sub, i)
}
rm(i, one)

# Some infections previously excluded might have slipped through, this is to make sure those are all removed. 


# Center the data around a pseudo peak time, with two weeks of extra negative samples on each side of the infection. 
id_list <- c(unique(model_dat$indid_inf[which(model_dat$infcluster >=1)]))
mod_dat <- data.frame()
model_dat$center_date <- NA
for(i in id_list) {
  sub <- model_dat %>%
    filter(indid_inf == i)
  tp <- c(sub$date[which(sub$CT == min(sub$CT))])
  sub$center_date <- sub$npsdatecol[which(sub$date == tp[1])]
  sub$date = sub$date - tp[1]
  sub <- sub %>%
    filter(date[which(date == min(date[which(CT != 37)]))] - date < 14 &
             date - date[which(date == max(date[which(CT != 37)]))] < 14)
  mod_dat <- rbind(mod_dat, sub)
}

rm(i, sub, tp, id_list)
```

## Save 
Finally save the two data frames for easier access for running the model.
```{r}
write.csv(flu_dat, "full_data.csv")
write.csv(mod_dat, "model_data.csv")
# REMAKE ONE DAT AFTER THESE EXCLUSIONS FOR ANALYSIS
in_model <- c(unique(mod_dat$indid_inf))
one_dat <- one_dat %>%
  filter(indid_inf %in% in_model)
rm(in_model)
write.csv(one_dat, "one_data.csv")
```


